---
title: "p8105_hw3_stl2137.Rmd"
output: github_document
---

# Homework 3
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
```

## Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```
To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. (INCOMPLETE)

The instacart dataset is comprised of `r nrow(instacart)` observations and `r ncol(instacart)` variables. 



```{r}
aisle_count <- instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```
There are a total of `count(aisle_count)` aisles, and the top 6 aisles where the most items are ordered from are fresh vegetables, fresh fruits, packaged vegetables fruits, yogurt, packcaged cheese, and water seltzer sparkling water. 

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart %>% 
  select(aisle, department) %>% 
  count(aisle) %>% 
  arrange(desc(n)) %>% 
  filter(n > 10000) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_col() +
  labs(
    x = "Name of Aisle",
    y = "Number of Items Ordered"
  ) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}

```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}

```


